Learning Features and Abstract Actions for Computing Generalized Plans

### 2018-11-12 doc/camera-ready/paper.tex
Blai edit

The paper is organized as follows. We provide the background  first and show then
how features and abstract actions can be learned by enforcing soundness and
completeness over a set of samples, from a pool of candidate features derived from the domain
predicates. The full computational model is then summarized, followed by experimental results
and a discussion.

@@

A solution to a generalized problem $\Q$ given the features $F$ and abstract
actions $A_F$ is a partial function $\pi$ that maps \emph{abstract states}
into \emph{abstract actions} such that $\pi$ solves all instances $P$ in $\Q$.
The plan or policy $\pi$ induces a trajectory $s_0,a_0,s_1, \ldots, s_n$ in
an instance $P$ of $\Q$ iff
1)~$s_0$ is the initial state in $P$,
2)~$a_i$ is one of the actions that instantiate $\pi(\bar{s}_i)$ in $s_i$,
3)~$a_i$ is applicable in $s_i$ and $s_{i+1}=f(a_i,s_i)$, and
4)~$s_n$ is a goal state of $P$, or $\pi(\bar{s}_n)$ is undefined,
not applicable in $\bar s_n$, or no applicable action $a_n$ in $P$ instantiates $\pi(\bar{s}_n)$.
The policy $\pi$ solves $P$ iff all trajectories induced by $\pi$ reach a goal state of $P$.

@@

We also consider an alternative SAT theory $T_G(\S,\F)$ that is similar to $T(\S,\F)$
but  smaller. It is obtained by marking  some of the transitions $(s,s')$ in $\S$  as  \emph{goal-relevant}.
Then, rather than creating abstract  actions to account for all the transitions, we only create
abstract actions to account for the marked transitions. This is achieved by drawing the states
$s$ and the transitions $(s,s')$ in formulas (\ref{eq:d1}) and (\ref{eq:d2}) from the
set of marked transitions in $\S$. The states $t$ and the transitions $(t,t')$, on the other hand,
are drawn from the sample set $\S$ exactly as before. This simplification preserves soundness over $\S$
but the completeness is not over $\S$ but over the marked transitions. 
The   transitions marked as goal relevant in $\S$  are obtained by computing one plan
for each sampled instance $P$ from $\Q$.

@@

\begin{enumerate}[{\small$\bullet$}]
  \item $C_A \leftarrow C_p, C_u, C_x$,  primitive, universal, nominals: $C_u$ denotes  universe,
   $C_x$ denotes $\set{x}$ for parameter $x$ if any,
  \item $C \leftarrow \neg C_A$, negation on primitive, universal, nominals,
  \item $C \leftarrow C \sqcap C'$, conjunctions,
  \item $C \leftarrow \exists R.C,\forall R.C$, first denotes $\{x:\!\exists y[R(x,y) \land C(y)]\}$, %  $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$;
    second $\{x:\forall y[R(x,y)\land C(y)]\}$, %,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
  \item $C \leftarrow R=R'$, denotes $\{ x : \forall y[R(x,y)=R'(x,y)]\}$, %$x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
  %  set of $y$'s  for which $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
  \item $R \leftarrow R_p, R_p^{-1}, R_p^+, [R_p^{-1}]^+$: primitive, inverse, and transitive closure of both.
%     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
%     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
 \end{enumerate}

@@

We have introduced a scheme for computing general plans that mixes
learning and planning: a \emph{learner} infers a general abstraction made
up of features and abstract actions by enforcing a notion of soundness
and completeness over a  set of samples, and a \emph{planner}
uses the abstraction, suitably transformed, to compute general plans. 
The number of samples required for obtaining correct general plans is
small, as the learner does not have to produce plans; it just
has to learn the relevant features for the planner to track.
Unlike purely learning approaches, the features and the policies
are transparent, and the scope and correctness
of the resulting general plans can be assessed.

@@

On the other hand, contexts involving non-deterministic actions may be
accommodated with a simple modification of the SAT theory.
Indeed, it would be enough to replace the formula (3) that links $D1$ and $D2$
with a formula that says that when $s$ and $t$ are indistinguishable
(i.e.\ $\neg D1(s,t)$), then for each action $a$ for $s$, there is an action $b$
for $t$ whose sets of possible outcomes are indistinguishable.

\subsubsection{Acknowledgments.}
H.\ Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
G.\ Franc\`{e}s is funded by the SSX project, European Research Council.
B.\ Bonet is partially funded by 2018 C\'atedra de Excelencia UC3M-Banco Santander award.

>>>>

The paper is organized as follows. We first provide background and then show 
how an abstract actions can be learned by enforcing soundness and
completeness over a set of samples from a pool of candidate features derived
from the domain predicates. The computational model is then summarized,
followed by experimental results and a discussion.

@@

A solution to a problem $\Q$ given the features $F$ and abstract
actions $A_F$ is a partial function $\pi$ that maps \emph{abstract states}
into \emph{abstract actions} such that $\pi$ solves all instances $P$ in $\Q$.
The plan or policy $\pi$ induces a trajectory $s_0,a_0,s_1, \ldots, s_n$ in
an instance $P$ of $\Q$ iff
1)~$s_0$ is the initial state in $P$,
2)~$a_i$ is one of the actions that instantiate $\pi(\bar{s}_i)$ in $s_i$,
3)~$a_i$ is applicable in $s_i$ and $s_{i+1}=f(a_i,s_i)$, and
4)~$s_n$ is a goal state of $P$, or $\pi(\bar{s}_n)$ is undefined,
not applicable in $\bar s_n$, or no applicable action $a_n$ in $P$ instantiates $\pi(\bar{s}_n)$.
The policy $\pi$ solves $P$ iff all trajectories induced by $\pi$ reach a goal state of $P$.

@@

We also consider an alternative SAT theory $T_G(\S,\F)$ that is similar to $T(\S,\F)$
but  smaller. It is obtained by marking  some transitions $(s,s')$ in $\S$  as  \emph{goal relevant}.
Then, rather than creating abstract  actions to account for all the transitions, we only create
abstract actions to account for the marked transitions. This is achieved by drawing the states
$s$ and the transitions $(s,s')$ in formulas \eqref{eq:d1} and \eqref{eq:d2} from the
set of marked transitions in $\S$. The states $t$ and the transitions $(t,t')$, on the other hand,
are drawn from the whole sample set $\S$ as before. This simplification preserves soundness over $\S$
but completeness is not over $\S$ but over the marked transitions in $\S$. 
The goal-relevant transitions in $\S$ are obtained by computing one plan
for each sampled instance $P$ from $\Q$.
%The transitions marked as goal relevant in $\S$  are obtained by computing one plan
%for each sampled instance $P$ from $\Q$.

@@

\begin{enumerate}[{\small$\bullet$}]
  \item $C_A \leftarrow C_p, C_u, C_x$,  primitive, universal, nominals: $C_u$ denotes  universe,
   $C_x$ denotes $\set{x}$ for parameter $x$ if any,
  \item $C \leftarrow \neg C_A$, negation on primitive, universal, nominals,
  \item $C \leftarrow C \sqcap C'$, conjunctions,
  \item $C \leftarrow \exists R.C,\forall R.C$, first denotes $\{x:\!\exists y[R(x,y) \land C(y)]\}$, %  $x$'s  s.t.\  $R(x,y) \land C(y)$ for some  $y$;
    the second denotes $\{x:\forall y[R(x,y)\land C(y)]\}$, %,  $x$'s  s.t. $R(x,y) \rightarrow C(y)$ for all $y$,
  \item $C \leftarrow R=R'$, denotes $\{ x : \forall y[R(x,y)=R'(x,y)]\}$, %$x$'s  s.t. $R(x,y)$ iff $R'(x,y)$ for all $y$,
  %  set of $y$'s  for which $R(x,y)$ equal to set of $y$'s s.t. $R'(x,y)$.
  \item $R \leftarrow R_p, R_p^{-1}, R_p^+, [R_p^{-1}]^+$: primitive, inverse, and transitive closure of both.
%     \item $R \leftarrow R \circ R'$, role composition: $(x,y)$ in $R \circ R'$ if $(x,z)$ in $R$ and $(z,y)$ in $R'$.
%     \item $R \leftarrow R:C$, denotes pairs $(x,y)$ in $R$ s.t. $C(y)$.
 \end{enumerate}

@@

We have introduced a scheme for computing general plans that mixes
learning and planning: a \emph{learner} infers an abstraction made
up of features and abstract actions by enforcing %a notion of
soundness and completeness over the samples, and a \emph{planner}
uses the abstraction, suitably transformed, to compute general plans. 
The number of samples required for obtaining correct general plans is
small, as the learner does not have to produce plans; it just
has to learn the relevant features for the planner to track.
Unlike purely learning approaches, the features and the policies
are transparent, and the scope and correctness
of the resulting general plans can be assessed.

@@

Finally, problems involving non-deterministic transitions may be
accommodated with a simple fix to the SAT theory.
It would be enough to replace the formula \eqref{eq:bridge1} that
links $D1$ and $D2$ with a formula that says that when $s$ and $t$ are
indistinguishable (i.e.\ $\neg D1(s,t)$), for each action $a$ on $s$,
there is an action $b$ on $t$ whose transitions are indistinguishable
from those of $a$ on $s$. For this, transitions $(s,s')$ in the sample
would need to be tagged with action labels as $(s,a,s')$.

%\bigskip
\subsection*{Acknowledgments}
B.\ Bonet is partially funded by 2018 C\'atedra de Excelencia UC3M-Banco Santander award.
G.\ Franc\`{e}s is funded by the SSX project, European Research Council.
H.\ Geffner is partially funded by grant TIN-2015-67959-P, MINECO, Spain.
